### 2025-08-06 cognitive-bias.qmd
20250806 - tweaks

Applied to fantasy football, consider that a Quarterback has had several lousy games in a row.
The gambler's fallacy might lead a manager to start the player under the belief that the player "is due" for a big game, expecting a strong performance from the player merely because they player has not had a good game in a while.

@@

The *anecdotal fallacy* is the error of drawing inferences based on personal experience or singular examples rather than based on stronger forms of evidence.
Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @sec-internalExternalValidityTradeoffs.
Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Wide Receiver by drafting with their first two picks, respectively.
The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite stronger evidence otherwise.

@@

The *narrative fallacy* is the error of creating a story with cause-and-effect explanations from random events.
According to @Kahneman2011, people find stories more compelling if they a) are simple (as opposed to complex), b) are concrete (as opposed to abstract), c) provide a greater attribution to talent, stupidity, and intentions (compared to luck), and d) focus on a few salient events that occured (rather than the many events that failed to occur).
For instance, consider that a player undergoes a challenging start to the season; however, they have higher-performing games in the middle of the season.
A fantasy football manager may construct a "causal narrative" that the player overcame obstacles and is now trending toward strong performance for the rest of the season, even though the stronger performance could reflect some natural and random fluctuation.

>>>>

Applied to fantasy football, consider that a Quarterback has had several lousy games in a row.
The gambler's fallacy might lead a manager to start the player under the belief that the player "is due" for a big game, expecting a strong performance from the player merely because the player has not had a good game in a while.

@@

The *anecdotal fallacy* is the error of drawing inferences based on personal experience or singular examples rather than based on stronger forms of evidence.
Although people tend to find stories and anecdotes compelling, they are among the weakest forms of evidence, as described in @sec-internalExternalValidityTradeoffs.
Consider that a fantasy football commentator writes an article about how they won their league by drafting a Quarterback and Tight End with their first two picks, respectively.
The anecdotal fallacy might lead a manager who reads the article to believe that they are more likely to win their league if they follow a similar strategy, despite stronger evidence otherwise.
It is important to be skeptical of anecdotes and personal experience.

@@

The *narrative fallacy* is the error of creating a story with cause-and-effect explanations from random events.
According to @Kahneman2011, people find stories more compelling if they a) are simple (as opposed to complex), b) are concrete (as opposed to abstract), c) provide a greater attribution to talent, stupidity, and intentions (compared to luck), and d) focus on a few salient events that occurred (rather than the many events that failed to occur).
For instance, consider that a player undergoes a challenging start to the season; however, they have higher-performing games in the middle of the season.
A fantasy football manager may construct a "causal narrative" that the player overcame obstacles and is now trending toward strong performance for the rest of the season, even though the stronger performance could reflect some natural and random fluctuation.
